---
title: "shenyi-logisticpca"
author: "Shenyi Jiang"
date: "2025-11-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
load("data/claims-clean.RData")
ls()
```

```{r}
# glimpse data cleaning
library(tidyverse)
glimpse(claims_clean)
names(claims_clean)
```


```{r}
library(logisticPCA)
table(claims_clean$bclass)

# change bclass to binary
claims_df <- claims_clean %>%
  mutate(
    bclass_bin = if_else(
      bclass == "Relevant claim content",
      1L,
      0L
    )
  )
table(claims_df$bclass_bin)
```

```{r}
# independent variable y (0/1)
y <- claims_df$bclass_bin

# matrix removes id bclass and bclass_bin
X <- claims_df %>%
  select(-.id, -bclass, -bclass_bin)
X_matrix <- as.matrix(X)
dim(X_matrix)
```

```{r}
# train/test set
set.seed(10162025)

n <- nrow(X_matrix)
train_idx <- sample(seq_len(n), size = floor(0.8*n))

X_train <- X_matrix[train_idx,]
X_test <- X_matrix[-train_idx,]

y_train <- y[train_idx]
y_test <- y[-train_idx]
```

```{r}
# dimension reduction
library(logisticPCA)

logpca_fit <- logisticPCA(X_train, k=10, m=6)
logpca_fit
```

```{r}
# train scores
train_scores <- logpca_fit$PCs
dim(train_scores)

test_scores <- predict(logpca_fit, X_test)
```

```{r}
# scores data frame
train_df <- as.data.frame(train_scores) %>%
  mutate(y=factor(y_train))

head(train_df)

# logistic regression
logit_mod <- glm(
  y ~ .,
  data = train_df,
  family = binomial
)
summary(logit_mod)
```

```{r}
# summarize
test_df <- as.data.frame(test_scores)

# predict probability
test_prob <- predict(logit_mod, newdata=test_df, type="response")
test_pred <- ifelse(test_prob >= 0.5, 1L, 0L)

# accuracy
mean(test_pred == y_test)
```
Using logistic principal component regression with 10 components, the model achieves an accuracy of 62.71% on the 20% hold-out test set.

## Summary
Task 2 — Logistic principal component regression
We used the cleaned bag-of-words representation of the claims webpages and applied logistic PCA to reduce the dimensionality of the word features. Using 10 logistic principal components as predictors in a logistic regression model for the binary label bclass (Relevant vs. Not Relevant), we obtained an accuracy of 62.71% on a hold-out test set (20% of the data). 
